Effectful Semantics for Substructural SSA

### 2023-11-08 pldi24/paper.tex
Tap

@@

\begin{definition}[Submonad] 
  A \textbf{submonad} \(\mb{S}\) of a monad \(\mb{T}\) on \(\ms{Set}\) is a collection of sets such that, for all \(A\), \(\mb{S}\;A \subseteq \mb{T}\;A\), \(\forall a \in A, \eta\;a \in \mb{S}\;A\), and, for all \(s \in \mb{S}\;A\), \(f: A \to \mb{S}\;B\), \(\mb{bind}\;s\;f \in \mb{S}\;A\). We say that a submonad of \(\mb{T}\) is \textbf{Elgot} if \(\mb{T}\) is Elgot and, for all \(f: A \to \mb{S}(B + A)\), \(f^\dagger: A \to \mb{S}\;B\) where \(f^\dagger\) is computed as in \(\mb{T}\).
\end{definition}
Some important examples of submonads include:
\begin{itemize}
  \item The \textbf{unit submonad} \(\ms{pure}(\mb{T})\) of every monad \(\mb{T}\) with \(\ms{pure}(\mb{T})\;A = \{\eta_A\;a \mid a \in A\}\); similarly, \(\mb{T}\) is itself a submonad of \(\mb{T}\).
  \item Given submonads \(\mb{S}_i\) of \(\mb{T}\), their elementwise intersections \(\bigwedge_i\mb{S}_i = \lambda A. \bigcap_i \mb{S}_i\;A\) are submonads, forming a meet-semilattice with bottom element \(\ms{pure}(\mb{T})\) and top element \(\mb{T}\)
  \item \todo{commuting absorbtive submonad of trace monad: example of...}
  \item \todo{central submonad. Go cite \cite{carette-central-2022}}
  \item \todo{deterministic, idempotent trace monad: relevant submonad}
  \item \todo{noneffectful trace monad: affine submonad}
  \item \todo{cap and centralize, cap and purify}
\end{itemize}
\TODO{in particular, these induce a substructural effectful category...}

>>>>

\begin{definition}[Effectful Category] \
  \TODO{this}
\end{definition}

\begin{definition}[(Symmetric) Premonoidal Subcategory]
  Given a (symmetric) premonoidal category \(\mc{C}\), we say a wide subcategory \(\mc{S} \subseteq \mc{C}\) is \textbf{(symmetric) premonoidal} if it is closed under whiskering and contains all associators, unitors, and symmetries. We say a subcategory is \textbf{monoidal} if, for all \(f: \mc{S}(A, B)\), \(g: \mc{S}(A', B')\), \(f \ltimes g = f \rtimes g\); we say it is \textbf{central} if this generalizes to the case where only one of \(f, g\) is in \(\mc{S}\) (i.e., if all morphisms in \(\mc{S}\) are central).
\end{definition} 
Note that the set of symmetric premonoidal subcategories, monoidal subcategories, and central subcategories of a category \(\mc{C}\) are all closed under intersections and therefore, in particular, form a complete meet-semilattice with shared bottom element \(\ms{Skeleton}(\mc{C})\), where \(\ms{Skeleton}(C)\) is the smallest possible symmetric premonodial subcategory. The largest symmetric premonoidal subcategory is of course \(\mc{C}\), while the largest central subcategory is the center of \(\mc{C}\) \(Z(\mb{C})\); the existence of a top-element means that in these cases we may induce the structure of a complete lattice in the usual manner.

\begin{definition}[Strictly Effectful Category] \
  \TODO{this}
\end{definition}

@@

\begin{definition}[Submonad] 
  A \textbf{submonad} \(\mb{S}\) of a monad \(\mb{T}\) on \(\ms{Set}\) is a collection of sets such that, for all \(A\), \(\mb{S}\;A \subseteq \mb{T}\;A\), \(\forall a \in A, \eta\;a \in \mb{S}\;A\), and, for all \(s \in \mb{S}\;A\), \(f: A \to \mb{S}\;B\), \(\mb{bind}\;s\;f \in \mb{S}\;A\). We say that a submonad of \(\mb{T}\) is \textbf{Elgot} if \(\mb{T}\) is Elgot and, for all \(f: A \to \mb{S}(B + A)\), \(f^\dagger: A \to \mb{S}\;B\) where \(f^\dagger\) is computed as in \(\mb{T}\).
\end{definition}
Some important examples of submonads include:
\begin{itemize}
  \item The \textbf{unit submonad} \(\ms{pure}(\mb{T})\) of every monad \(\mb{T}\) with \(\ms{pure}(\mb{T})\;A = \{\eta_A\;a \mid a \in A\}\); similarly, \(\mb{T}\) is itself a submonad of \(\mb{T}\).
  \item Given submonads \(\mb{S}_i\) of \(\mb{T}\), their elementwise intersections \(\bigwedge_i\mb{S}_i = \lambda A. \bigcap_i \mb{S}_i\;A\) are submonads, forming a complete meet-semilattice with bottom element \(\ms{pure}(\mb{T})\) and top element \(\mb{T}\); this induces a complete lattice in the standard manner
  \item \todo{commuting absorbtive submonad of trace monad: example of...}
  \item \todo{central submonad. Go cite \cite{carette-central-2022}}
  \item \todo{deterministic, idempotent trace monad: relevant submonad}
  \item \todo{noneffectful trace monad: affine submonad}
  \item \todo{cap and centralize, cap and purify}
\end{itemize}
\TODO{a submonad induces a wide subcategory}

\TODO{in particular, these induce a substructural effectful category...}
